### 2013-01-30 Docs/Thesis/system_design_and_interface.tex
ANH thesis revisions

\subsection{Evaluation Monad}
\label{section:System_Design_and_Interface-Types-Evaluation_Monad}

A {\em monad} is a standard, composable abstraction for writing functions with
a context, used in Haskell for IO~\cite{PeytonJones1993,PeytonJones2001} and
other tasks. A monad is simply a 1-arity type constructor together with two
functions. The first function, {\tt return}, takes a value of type {\tt a} and
produces a value of type {\tt m a}. The second, called {\tt bind} and stylized
in the Haskell standard library as the infix operator {\tt (>>=)}, takes a value
of type {\tt m a} and a function of type {\tt a -> m b} and produces a value of
type {\tt m b}. This allows a value to be operated on out of the context and a
new context to be assigned.

A monad can have other primitives which manipulate the context in some way. For 
instance, the primtives in Haskell's {\tt IO} monad produce actions which, when
interpreted as part of the {\tt main} action, produce some side-effect. The
{\tt State} monad provides {\tt get} and {\tt put} operations to work with a 
state value stored in the context.

Monad transformers~\cite{Jones1995} provide a means to combine the functionality
of multiple monads. A monad transformer is a monad with an extra type parameter.
This type parameter is instantiated with the type constructor of the underlying
monad, and an extra operation {\tt lift} is provided which converts values in
the underlying monad to values in the monad transformer.

In TimeFlies, the evaluation monad is a monad transformer. This permits it to be
used in conjunction with the {\tt IO} monad (or any other monad) to describe how
input is to be obtained for the signal function being evaluated, and how outputs
are to be handled.

@@

The type of the evaluation monad must track the input and output type of the
signal function. The monad is a state monad, and its state is a record with
fields for a mapping from outputs to handling actions, the current state of the
signal function, the last time the signal function was sampled, and the
currently-unsampled input. There are thus four type parameters to the monad's
type constructor: the input signal vector (to type the signal function and
unsampled input), the output signal vector (to type the signal function and
output handlers), the type of the underlying monad, and the monadic type
parameter. The type is shown in Figure~\ref{figure:evaluation_monad_types}.

>>>>

\subsection{Evaluation Monad Transformer}
\label{section:System_Design_and_Interface-Types-Evaluation_Monad_Transformer}

TimeFlies provides a monad transformer for evaluating signal functions. Thi
monad transformer may be used in conjunction with the {\tt IO} monad (or any
other monad) to describe how input is to be obtained for the signal function
being evaluated, and how outputs are to be handled.

@@

The type of the evaluation monad must track the input and output type of the
signal function. The monad is a state monad, and its state is a record (the
{\tt SFEvalState} type) with fields for a mapping from outputs to handling
actions, the current state of the signal function, the last time the signal
function was sampled, and the currently-unsampled input. There are thus four
type parameters to the monad's type constructor: the input signal vector (to
type the signal function and unsampled input), the output signal vector (to type
the signal function and output handlers), the type of the underlying monad, and
the monadic type parameter. The type is shown in
Figure~\ref{figure:evaluation_monad_types}.
